### 2015-04-14 src/paper1/paper1.tex
Fixed minor typos

\subsection{Model validation: independent test of neck noise}

@@

The SBT experiment, performed in seven subjects, tested the accuracy and precision of SBT percepts, near upright (SBT0), at \siang{45} and \siang{90} right side down (\SBT{45} and \SBT{90}), and at \siang{45} and \siang{90} left side down (\SBT{-45} and \SBT{-90}). The same subjects also performed the SVV experiment, tested at nine roll-tilt angles, ranging from \siang{-120} to \siang{120} in \siang{30} intervals. \figref{p1:fig3} shows the results of a typical subject (S1) in both tasks. The top panels show the proportion of CW responses for the five SBT tasks, relative to the reference orientation. For an ideal observer, all psychometric functions would resemble a step centred at zero. Across the five reference orientations (\siang{0}, \siang{{\textpm}45}, or \siang{{\textpm}90}), the psychometric data indicate underestimations and overestimations of perceived body angle, but no consistent bias, which resembles previous reports \cite{mittelstaedt1983, mast1996, jarchow1999, vanbeuzekom2001} that body-tilt perception is accurate on average. We fitted psychometric curves through these data (see \nameref{p1:sec:methods}, Eq. \ref{p1:eqn1}), to obtain estimates for the mean ($\mu$), SD ($\sigma$), and lapse rate ($\lambda$). Parameter $\mu$ is a measure for the accuracy of the subject's body-tilt percept. Perceptual variability, inversely related to precision, is reflected by $\sigma^2$, whereas the lapse rate ($\lambda = 0.06$) accounts for stimulus-independent errors \cite{wichmann2001}. In all five SBT tasks, the $\mu$ values are relatively close to the veridical reference orientation (0\textdegree, 45\textdegree, or 90\textdegree), i.e., errors are \textless5\textdegree. The psychometric fits further show that variability is lower in the \SBT{0} task, with $\sigma \approx 4\degree$, than in the \SBT{\textpm45} and \SBT{\textpm90} task, where $\sigma \approx 10\degree$.

@@

To further validate the model, we independently tested one of its predictions that can be assessed experimentally in isolation: head-on-body variance. In supine position, subjects judged their head orientation (CW/CCW) relative to the body midline after it had been passively roll-rotated with speeds subthreshold for the canals to various angles (see \nameref{p1:sec:methods}, Model validation: independent test of neck noise). Psychometric fits indicate no systematic bias in these head-on-body percepts (data not shown). Figure \ref{p1:fig6} depicts the experimental noise levels derived from these psychometric fits and the predicted values provided by the model, including their 95\% confidence intervals. Because the variance of the estimates increases with the average head-on-body percept, we performed a regression on the log-transformed data \cite{hopkins2000}. The significant correlation between predicted and measured neck noise levels (slope, 1.03; intercept, -0.1; $p = 0.04$) provides an independent confirmation of the proposed model.

@@

\begin{figure}
    \includegraphics[width=0.80\textwidth]{src/paper1/figure8.pdf}
    \caption{Bayesian computations in single and multiple trials. A-C, Single trial. D-F, Multiple trials. G-I, Resulting distributions. For further explanation, see \nameref{p1:sec:appendix}.}
    \label{p1:fig8}
\end{figure}

>>>>

\subsection{Model validation: independent test of neck noise}
\label{p1:sec:model:necknoise}

@@

The SBT experiment, performed in seven subjects, tested the accuracy and precision of SBT percepts, near upright (SBT0), at \siang{45} and \siang{90} right side down (\SBT{45} and \SBT{90}), and at \siang{45} and \siang{90} left side down (\SBT{-45} and \SBT{-90}). The same subjects also performed the SVV experiment, tested at nine roll-tilt angles, ranging from \siang{-120} to \siang{120} in \siang{30} intervals. \figref{p1:fig3} shows the results of a typical subject (S1) in both tasks. The top panels show the proportion of CW responses for the five SBT tasks, relative to the reference orientation. For an ideal observer, all psychometric functions would resemble a step centred at zero. Across the five reference orientations (\siang{0}, \siang{{\textpm}45}, or \siang{{\textpm}90}), the psychometric data indicate underestimations and overestimations of perceived body angle, but no consistent bias, which resembles previous reports \cite{mittelstaedt1983, mast1996, jarchow1999, vanbeuzekom2001} that body-tilt perception is accurate on average. We fitted psychometric curves through these data (see \nameref{p1:sec:methods}, \eqnref{p1:eqn1}), to obtain estimates for the mean ($\mu$), SD ($\sigma$), and lapse rate ($\lambda$). Parameter $\mu$ is a measure for the accuracy of the subject's body-tilt percept. Perceptual variability, inversely related to precision, is reflected by $\sigma^2$, whereas the lapse rate ($\lambda = 0.06$) accounts for stimulus-independent errors \cite{wichmann2001}. In all five SBT tasks, the $\mu$ values are relatively close to the veridical reference orientation (0\textdegree, 45\textdegree, or 90\textdegree), i.e., errors are \textless5\textdegree. The psychometric fits further show that variability is lower in the \SBT{0} task, with $\sigma \approx 4\degree$, than in the \SBT{\textpm45} and \SBT{\textpm90} task, where $\sigma \approx 10\degree$.

@@

To further validate the model, we independently tested one of its predictions that can be assessed experimentally in isolation: head-on-body variance. In supine position, subjects judged their head orientation (CW/CCW) relative to the body midline after it had been passively roll-rotated with speeds subthreshold for the canals to various angles (see \nameref{p1:sec:model:necknoise}). Psychometric fits indicate no systematic bias in these head-on-body percepts (data not shown). Figure \ref{p1:fig6} depicts the experimental noise levels derived from these psychometric fits and the predicted values provided by the model, including their 95\% confidence intervals. Because the variance of the estimates increases with the average head-on-body percept, we performed a regression on the log-transformed data \cite{hopkins2000}. The significant correlation between predicted and measured neck noise levels (slope, 1.03; intercept, -0.1; $p = 0.04$) provides an independent confirmation of the proposed model.

@@

\begin{figure}
    \includegraphics[width=0.80\textwidth]{src/paper1/figure8.pdf}
    \caption{Bayesian computations in single and multiple trials. \panel{A-C}, Single trial. \panel{D-F}, Multiple trials. \panel{G-I}, Resulting distributions. For further explanation, see \nameref{p1:sec:appendix}.}
    \label{p1:fig8}
\end{figure}

### 2015-04-14 src/paper3/paper3.tex
Fixed minor typos

\small {\bf Abstract} Self-motion is typically accompanied by compensatory eye movements that help minimize retinal slip and maximize dynamic visual acuity. To date, it is unknown whether these eye movements also have a reversed role, serving as a cue for self-motion perception. To address this question, we had participants ($n=8$) judge self-motion during different eye movement conditions in the absence of full-field optic flow.  In a 2-AFC task, participants indicated whether the second of two successive passive lateral whole-body translations was longer or shorter than the first. Eye movements during each translation were world-stationary, or body-stationary in an otherwise dark room. Results of these two conditions show that the perceived translations were shorter with body-fixed gaze compared to world-fixed gaze. Using a linear model, we estimated the relative contributions of vestibular and eye movement signals to self-motion perception and found that eye movement signals contribute approximately 25 percent. The model was independently validated by successfully predicting the effects of eye movements on self-motion in a third condition without any visual fixation, i.e. when the eyes were free to move. We conclude that eye movement signals influence self-motion perception, even in the absence of visual stimulation, and even when oculomotor and vestibular estimates are in conflict, e.g. during body-fixed gaze. We hypothesize that adverse consequences of this seemingly inflexible arrangement are minimal under natural conditions because eye movements and self-motion are highly correlated, and because eye movements are most often accompanied by veridical optic flow cues to self-motion.

@@

An accurate estimate of self-motion is important to guide interactions with the environment. During passive self-motion both vestibular and optic flow signals provide information about self-motion \cite{gibson1955, benson1986, harris2000, israel1989, angelaki2005, carriot2013, chen2010}. However, also compensatory eye movements that maintain fixation on world-fixed objects carry self-motion information. These eye movements are driven by retinal slip or vestibular signals. For example, the linear vestibulo-ocular reflex (LVOR) stabilizes gaze during head translations, even in complete darkness \cite{paige1989,medendorp2002,angelaki2004}.  Many studies have shown that the brain uses oculomotor signals to extract the optic flow component related to self-motion \cite{warren1988, royden1992, freeman1998, lappe1999}, but to our knowledge a direct influence of eye movements on self-motion perception has not been investigated. Here we study whether these oculomotor signals are also used to estimate self-motion directly.

@@

To test whether eye movements are used in self-motion perception, we employed a two-alternative forced choice (2-AFC) paradigm in which participants were presented with two consecutive lateral translations. They had to indicate whether the second translation was longer or shorter than the first. Eye movements during each interval were either constrained using a  body or world stationary fixation point or not constrained at all (i.e. free). We show that identical translations were perceived shorter when gaze was body stationary compared to world stationary. Furthermore, using a linear model we predicted perceived displacement during the free gaze condition based on vestibular signals and unconstrained eye movements. We conclude that eye movements influence self-motion perception even in the absence of optic flow or other visual stimulation.

@@

A motorized linear sled \cite<see>[for details]{clemens2012} was used to laterally translate participants following a minimum jerk profile \cite{flash1985} of fixed duration (1 \si{\second}) and amplitudes ranging from 1 to 27 \si{\centi\metre}. Participants were seated on the sled such that the inter-aural axis aligned with the motion axis. They were restrained using a five-point seat belt and a chin rest. In addition, the head was held in place using a sled-fixed mold which resembled head-phones and pressed down on the head surrounding the pinnae. Auditory cues were suppressed using white noise presented through in-ear headphones. Experiments were conducted in complete darkness except for visual fixation points, projected by a laser pointer on a black bar 50 \si{\centi\metre} in front of the participant at eye level. Laser pointers used to project body-fixed targets were attached to the sled. Those used to project world-fixed targets were mounted on the wall behind the sled.

@@

\subsection{Paradigm}
We used a two-alternative forced choice (2-AFC) task to measure perceived linear self-motion across three different eye fixation types: world-fixed, body-fixed, and unconstrained (free) fixation. We refer to these as world, body and free, respectively. A trial contained two sequential translation intervals of equal duration (1 \si{\second}) and in the same direction (either leftward or rightward). Different fixation types were presented in the two translation intervals. Participants were instructed to judge whether the translation during the second interval was longer or shorter compared to the first interval. They were additionally instructed to always look at the fixation point when it was visible; no instructions were given for when the fixation point was switched off (i.e. during free fixation).

@@

    \caption{\panelref{A} Time course of key events within a single trial. In each of the two intervals, a 0.50 \si{\second} fixation period (red) precedes the lateral translation (yellow). A 1.75 \si{\second} long delay period (shown in white) separates the two intervals. After the second translation, the participant responded whether this second translation was longer or shorter than the first. \panelref{B} Top-view illustrating key events during a body vs. world trial. First panel: participant fixates the world-fixed target (red cross) at the start of the first interval. Second panel: translation with world-fixed fixation target. Third panel: body-fixed fixation at start of second fixation interval. Fourth panel: translation with body-fixed fixation in second interval.}
    \label{p3:fig1}
\end{figure}

@@

Thus, a trial consists of two translations with different fixation types; in the three main conditions we compare the body versus world, world versus free, and body versus free fixation types. For each main condition, we varied which fixation type served as the reference stimulus and the order in which reference and probe were presented, which gives a total of four variations per main condition (see \tabref{p3:tab1}). In addition we varied translation direction (either leftward or rightward on consecutive trials). The amplitude of the probe translation was adaptively chosen using the Psi method. This method picks the amplitude for the next trial which maximizes the expected decrease in entropy based on participants' responses to earlier trials \cite{kontsevich1999}. This was done separately for all 24 trial types (3 main conditions x 2 reference stimuli x 2 reference/probe orders x 2 translation directions; see \tabref{p3:tab1}). A total of 25 trials were collected per trial type yielding a total of 200 trials for each of the three main conditions.

@@

For each combination of the three main conditions, and  the two reference/probe orders (see \tabref{p3:tab1}), we quantified the perceived probe translation  by calculating the probability of the probe translation judged longer compared to the 10 \si{\centi\metre} reference translation as a function of actual probe translation, given by x. We used a maximum likelihood fit of a cumulative Gaussian function to summarize the psychometric data:

@@

Using a simple cue integration model, we investigated whether inter-subject and inter-condition differences in the observed PSEs in conditions containing a translation under free fixation depend on actual eye movement behavior. We modeled perceived distance, $p$, as a weighted linear combination of a vestibular and an oculomotor estimate of translation (\eqnref{p3:eq2}). We assumed that the vestibular estimate is equal to the actual translation, $m$, and that the oculomotor estimate is equal to expected eye movement given the  actual, $\hat{\varphi}_{m_i}$. As the weights represent the relative contributions of the oculomotor and vestibular systems, they can sum to any arbitrary value; in \eqnref{p3:eq2} their sum is fixed to 1. Thus, the weighting parameter $\alpha$ regulates the eye movement contribution and $1 - \alpha$  the vestibular contribution:

@@

In the present experiment, the reference displacement, $m_r$, was always 10 \si{\centi\metre} and the probe displacement, $m_p$, was equal to the measured PSE for the presented combination of fixation types (i.e, $PSE$ in \eqnref{p3:eq1}). This model (i.e. \eqnref{p3:eq3}) was then fit to data from the body and world conditions using linear regression, finding weight α that minimizes the sum of squared errors ($\sum{\epsilon^2}$).

@@

    \caption{Psychometric curves (colored lines) and associated binned data (circles) for one participant (top row). Circle size represents the number of trials within each 2 \si{\centi\metre} bin. Binning was only done in order to visualize this participant's responses and was not used otherwise. Gray lines show psychometric curves before collapsing across reference order. \panel{A} Body-world comparison; body reference, dark red; world reference, light red. \panel{B} World-free comparison; world reference, light green; free reference, dark green. \panel{C}  Body-free comparison; body reference, dark blue; free reference, light blue. \newline
PSEs for all participants and the average {\textpm}SE (bottom row). \panel{D} Body-world (dark red) and world-body (light red) conditions. \panel{E} World-free (light green) and free-world (dark green) conditions. \panel{F} Body-free (dark blue) and free-body (light blue) conditions. Because a t-test revealed a main effect of reference order, \ttest{47}{-5.2}{0.01}, we used the mean PSE across reference order (e.g. \figref{p3:fig3}, gray lines) instead of the PSE collapsing across reference order (e.g. \figref{p3:fig3}, colored lines); these values were not significantly different.}
    \label{p3:fig2}
\end{figure}

The performance of one participant is illustrated in the left column of \figref{p3:fig2}. Each row shows one main condition: body versus world fixation (top/red), world versus free fixation (middle/green), and body versus free fixation (bottom/blue). The lighter and darker colors in each panel indicate which fixation type was the reference movement (see Legend). The shift of the psychometric functions relative to the 10 \si{\centi\metre} reference (i.e. the PSE) quantifies the influence of fixation type. For example, the rightward shift of the light red curve in \panelref{p3:fig2}{A} means that for a body fixation a longer translation (\about 19 \si{\centi\metre}) was required for that translation to be perceived equivalent to a 10\si{\centi\metre} reference translation with world fixation. On the other hand, the leftward shift of the dark red curve means that a shorter translation with world fixation (\about 7 \si{\centi\metre}) was required for that translation to be perceived equivalent to the 10 \si{\centi\metre} reference translation with body fixation. Together, these oppositely directed shifts demonstrate that translations with world fixation were perceived longer than equivalent translations with body fixation, regardless of which translation was the reference.  Similarly, the shifts in \panelref{p3:fig2}{B} shows that world fixation translations were also perceived to be longer than free fixation movements and \panelref{p3:fig2}{C} shows that free fixation translations were perceived to be longer than body fixation translations. Note that \figref{p3:fig2} also shows effects on slope, which will be further discussed in the paragraph \nameref{p3:sec:precision}.

@@

    \caption{\panel{A} Actual (solid lines) and ideal (dashed lines) eye movement traces of one participant during world fixation (purple), body fixation (brown), and free fixation (black). All traces shown are for 10 \si{\centi\metre} reference movements. \panel{B} Normalized eye position for each participant (\textpm95\% confidence interval) at the end of translation interval (error bars) for world fixation (purple), body fixation (brown) and free fixation (blue). In addition, the average {\textpm}SE across all participants is shown. Zero indicates that the eyes remained stationary relative to the body, and one indicates that eye position was perfectly world-fixed.}
    \label{p3:fig3}
\end{figure}

@@

In order to relate psychophysical performance to eye movement behavior we recorded and analyzed eye movements during both intervals of every trial for all subjects. Exemplar eye traces for the 10 \si{\centi\metre} reference translation for the three fixation types are depicted in \panelref{p3:fig3}{A}. Fixation behavior was quite accurate for both body fixations, where no eye movements were expected, and world fixation, where eye movement excursions of \siang{11} were expected, seemingly supported by  catch-up saccades. Under free fixation, the amount of eye movement was intermediate between body and world fixation and behavior was more variable. A similar pattern was observed in all participants, as illustrated by the normalized eye movement data (see \nameref{p3:sec:methods}, and \panelref{p3:fig3}{B}).

@@

    \caption{Eye movement based prediction for the PSE plotted against the actual PSE.  A data point (symbol) is shown for each participant (symbol shape) and condition (symbol color) pair, following the same color scheme as in \figref{p3:fig2}. The identity line, corresponding to a perfect prediction, is shown in black.}
    \label{p3:fig4}
\end{figure}

@@

    \caption{Estimated eye movement contribution (α) to the perception of self-motion (see Equation 5). Standard deviations are based on a bootstrap for each participant.}

@@

To quantify the role of eye movements in self-motion perception, we tested a linear model (see \nameref{p3:sec:model}) in which perceived translation is a weighted average of a vestibular estimate (equal to the actual translation) and an oculomotor estimate (equal to the normalized eye movement times the actual translation; \eqnref{p3:eq2}). This model contains a single free parameter ($\alpha$), which corresponds to the relative weight given to the oculomotor estimate. We fit this model to the two body versus world conditions and obtained the value of the oculomotor weight for every subject (\tabref{p3:tab2}). The average oculomotor weight is 0.25 \textpm0.12 (SD), indicating that the contribution of the eye movement signal to the self-motion estimate is about 25 percent. Note that participant 4, whose oculomotor weight is furthest from this mean ($\alpha = 0.06$), also shows a radically different eye movement gain during the free-fixation (see \figref{p3:fig3}B). We then used these oculomotor weights along with the normalized eye movement values to predict the PSEs in the remaining four conditions according to \eqnref{p3:eq5}. The predicted PSEs are plotted against the actually observed PSEs in \figref{p3:fig4}. The positive correlation ($\rho = 0.78$, $p < 0.01$) between observed and predicted PSEs suggests that eye movements are indeed used in self-motion perception, even in the absence of a fixation point (i.e. during free fixation). Furthermore, the fact that data points generally cluster near the unity line shows that our simple model does reasonably well in predicting perceptual performance across subjects and conditions based on oculomotor weight and normalized eye movement magnitude only. This holds true even for subject 7 whose oculomotor weight (\tabref{p3:tab2}) was approximately double the average, yet whose data points remain close to the unity line.

@@

    \caption{Effect of difference movement amplitude between reference and probe interval (i.e. the bias) on response uncertainty ($\sigma$). A data point is shown for every participant and condition (symbol color) pair, following the same color scheme as in \figref{p3:fig2}. The dashed black line is the linear regression trend line.}

@@

 To quantitatively characterize the separate vestibular and eye movement contributions, we fit a single parameter model to the perceptual responses for the body versus world comparison conditions and validated this model independently by predicting the effects of eye movements on self-motion perception during free fixation conditions. This model takes into account subject specific oculomotor weight and eye movement patterns. Based on these inputs it accurately predicts the responses in the free fixation conditions. This demonstrates that extra-retinal eye movement signals are used as a cue in the perception of self-motion, contributing significantly to the self-motion percept with a weight of approximately 25 percent, even in the absence of optic flow.

It is surprising that an influence of eye movements can be observed even for body- stationary fixations, during which the stationary eye movement signal is clearly in conflict with the non-zero vestibular signal. While this demonstrates the strength of the assumption that fixation targets are world-stationary, it raises the question how reliable this assumption is. Simultaneous recording of angular head and eye movements during natural behavior reveals that approximately 80 percent of eye movements can be classified as compensatory, i.e. eye movements directed opposite to head movement and therefore consistent with maintenance of world-fixed fixation \cite{einhauser2007}. Similarly, other studies have shown that world-stationary fixations are common for many every day activities, ranging from making a cup of tea \cite{hayhoe2014} to driving a car \cite{land1994}, to walking \cite{foulsham2011} and even reaching, where people tend to look at the source and destination of the object, but not at the hand \cite{flanagan2003}. Because world-stationary fixations are so common, the natural world statistics imply that self-motion and eye movements are highly correlated, thus making eye movements a fairly reliable cue for self-motion.

@@

Studies conducted to characterize vestibular-only sensitivity are often performed in complete darkness or with closed eyes \cite{grabherr2008,  macneilage2010b, macneilage2010a, roditi2012, valko2012, nesti2014}. However, the results of our free-fixation condition suggest that even under these circumstances, results could easily be influenced by vestibularly driven eye movements. Overall, we suggest that any study concerned with self-motion processing must consider the possible influence of eye movements.

>>>>

\small {\bf Abstract} Self-motion is typically accompanied by compensatory eye movements that help minimise retinal slip and maximise dynamic visual acuity. To date, it is unknown whether these eye movements also have a reversed role, serving as a cue for self-motion perception. To address this question, we had participants ($n=8$) judge self-motion during different eye movement conditions in the absence of full-field optic flow.  In a 2AFC task, participants indicated whether the second of two successive passive lateral whole-body translations was longer or shorter than the first. Eye movements during each translation were world-stationary, or body-stationary in an otherwise dark room. Results of these two conditions show that the perceived translations were shorter with body-fixed gaze compared to world-fixed gaze. Using a linear model, we estimated the relative contributions of vestibular and eye movement signals to self-motion perception and found that eye movement signals contribute approximately 25 percent. The model was independently validated by successfully predicting the effects of eye movements on self-motion in a third condition without any visual fixation, i.e. when the eyes were free to move. We conclude that eye movement signals influence self-motion perception, even in the absence of visual stimulation, and even when oculomotor and vestibular estimates are in conflict, e.g. during body-fixed gaze. We hypothesise that adverse consequences of this seemingly inflexible arrangement are minimal under natural conditions because eye movements and self-motion are highly correlated, and because eye movements are most often accompanied by veridical optic flow cues to self-motion.

@@

An accurate estimate of self-motion is important to guide interactions with the environment. During passive self-motion both vestibular and optic flow signals provide information about self-motion \cite{gibson1955, benson1986, harris2000, israel1989, angelaki2005, carriot2013, chen2010}. However, also compensatory eye movements that maintain fixation on world-fixed objects carry self-motion information. These eye movements are driven by retinal slip or vestibular signals. For example, the linear vestibulo-ocular reflex (LVOR) stabilises gaze during head translations, even in complete darkness \cite{paige1989,medendorp2002,angelaki2004}.  Many studies have shown that the brain uses oculomotor signals to extract the optic flow component related to self-motion \cite{warren1988, royden1992, freeman1998, lappe1999}, but to our knowledge a direct influence of eye movements on self-motion perception has not been investigated. Here we study whether these oculomotor signals are also used to estimate self-motion directly.

@@

To test whether eye movements are used in self-motion perception, we employed a two-alternative forced choice (2AFC) paradigm in which participants were presented with two consecutive lateral translations. They had to indicate whether the second translation was longer or shorter than the first. Eye movements during each interval were either constrained using a  body or world stationary fixation point or not constrained at all (i.e. free). We show that identical translations were perceived shorter when gaze was body stationary compared to world stationary. Furthermore, using a linear model we predicted perceived displacement during the free gaze condition based on vestibular signals and unconstrained eye movements. We conclude that eye movements influence self-motion perception even in the absence of optic flow or other visual stimulation.

@@

A motorised linear sled \cite<see>[for details]{clemens2012} was used to laterally translate participants following a minimum jerk profile \cite{flash1985} of fixed duration (1 \si{\second}) and amplitudes ranging from 1 to 27 \si{\centi\metre}. Participants were seated on the sled such that the inter-aural axis aligned with the motion axis. They were restrained using a five-point seat belt and a chin rest. In addition, the head was held in place using a sled-fixed mold which resembled head-phones and pressed down on the head surrounding the pinnae. Auditory cues were suppressed using white noise presented through in-ear headphones. Experiments were conducted in complete darkness except for visual fixation points, projected by a laser pointer on a black bar 50 \si{\centi\metre} in front of the participant at eye level. Laser pointers used to project body-fixed targets were attached to the sled. Those used to project world-fixed targets were mounted on the wall behind the sled.

@@

\subsection{Paradigm}
We used a two-alternative forced choice (2AFC) task to measure perceived linear self-motion across three different eye fixation types: world-fixed, body-fixed, and unconstrained (free) fixation. We refer to these as world, body and free, respectively. A trial contained two sequential translation intervals of equal duration (1 \si{\second}) and in the same direction (either leftward or rightward). Different fixation types were presented in the two translation intervals. Participants were instructed to judge whether the translation during the second interval was longer or shorter compared to the first interval. They were additionally instructed to always look at the fixation point when it was visible; no instructions were given for when the fixation point was switched off (i.e. during free fixation).

@@

    \caption{\panel{A} Time course of key events within a single trial. In each of the two intervals, a 0.50 \si{\second} fixation period (red) precedes the lateral translation (yellow). A 1.75 \si{\second} long delay period (shown in white) separates the two intervals. After the second translation, the participant responded whether this second translation was longer or shorter than the first. \panel{B} Top-view illustrating key events during a body vs. world trial. First panel: participant fixates the world-fixed target (red cross) at the start of the first interval. Second panel: translation with world-fixed fixation target. Third panel: body-fixed fixation at start of second fixation interval. Fourth panel: translation with body-fixed fixation in second interval.}
    \label{p3:fig1}
\end{figure}

@@

Thus, a trial consists of two translations with different fixation types; in the three main conditions we compare the body versus world, world versus free, and body versus free fixation types. For each main condition, we varied which fixation type served as the reference stimulus and the order in which reference and probe were presented, which gives a total of four variations per main condition (see \tabref{p3:tab1}). In addition we varied translation direction (either leftward or rightward on consecutive trials). The amplitude of the probe translation was adaptively chosen using the Psi method. This method picks the amplitude for the next trial which maximises the expected decrease in entropy based on participants' responses to earlier trials \cite{kontsevich1999}. This was done separately for all 24 trial types (3 main conditions x 2 reference stimuli x 2 reference/probe orders x 2 translation directions; see \tabref{p3:tab1}). A total of 25 trials were collected per trial type yielding a total of 200 trials for each of the three main conditions.

@@

For each combination of the three main conditions, and  the two reference/probe orders (see \tabref{p3:tab1}), we quantified the perceived probe translation  by calculating the probability of the probe translation judged longer compared to the 10 \si{\centi\metre} reference translation as a function of actual probe translation, given by $x$. We used a maximum likelihood fit of a cumulative Gaussian function to summarise the psychometric data:

@@

Using a simple cue integration model, we investigated whether inter-subject and inter-condition differences in the observed PSEs in conditions containing a translation under free fixation depend on actual eye movement behaviour. We modelled perceived distance, $p$, as a weighted linear combination of a vestibular and an oculomotor estimate of translation (\eqnref{p3:eq2}). We assumed that the vestibular estimate is equal to the actual translation, $m$, and that the oculomotor estimate is equal to expected eye movement given the  actual, $\hat{\varphi}_{m_i}$. As the weights represent the relative contributions of the oculomotor and vestibular systems, they can sum to any arbitrary value; in \eqnref{p3:eq2} their sum is fixed to 1. Thus, the weighting parameter $\alpha$ regulates the eye movement contribution and $1 - \alpha$  the vestibular contribution:

@@

In the present experiment, the reference displacement, $m_r$, was always 10 \si{\centi\metre} and the probe displacement, $m_p$, was equal to the measured PSE for the presented combination of fixation types (i.e, $PSE$ in \eqnref{p3:eq1}). This model (i.e. \eqnref{p3:eq3}) was then fit to data from the body and world conditions using linear regression, finding weight α that minimises the sum of squared errors ($\sum{\epsilon^2}$).

@@

    \caption{Psychometric curves (coloured lines) and associated binned data (circles) for one participant (top row). Circle size represents the number of trials within each 2 \si{\centi\metre} bin. Binning was only done in order to visualise this participant's responses and was not used otherwise. Gray lines show psychometric curves before collapsing across reference order. \panel{A} Body-world comparison; body reference, dark red; world reference, light red. \panel{B} World-free comparison; world reference, light green; free reference, dark green. \panel{C}  Body-free comparison; body reference, dark blue; free reference, light blue. \newline
PSEs for all participants and the average {\textpm}SE (bottom row). \panel{D} Body-world (dark red) and world-body (light red) conditions. \panel{E} World-free (light green) and free-world (dark green) conditions. \panel{F} Body-free (dark blue) and free-body (light blue) conditions. Because a t-test revealed a main effect of reference order, \ttest{47}{-5.2}{0.01}, we used the mean PSE across reference order (e.g. \figref{p3:fig3}, gray lines) instead of the PSE collapsing across reference order (e.g. \figref{p3:fig3}, coloured lines); these values were not significantly different.}
    \label{p3:fig2}
\end{figure}

The performance of one participant is illustrated in the left column of \figref{p3:fig2}. Each row shows one main condition: body versus world fixation (top/red), world versus free fixation (middle/green), and body versus free fixation (bottom/blue). The lighter and darker colours in each panel indicate which fixation type was the reference movement (see Legend). The shift of the psychometric functions relative to the 10 \si{\centi\metre} reference (i.e. the PSE) quantifies the influence of fixation type. For example, the rightward shift of the light red curve in \panelref{p3:fig2}{A} means that for a body fixation a longer translation (\about 19 \si{\centi\metre}) was required for that translation to be perceived equivalent to a 10\si{\centi\metre} reference translation with world fixation. On the other hand, the leftward shift of the dark red curve means that a shorter translation with world fixation (\about 7 \si{\centi\metre}) was required for that translation to be perceived equivalent to the 10 \si{\centi\metre} reference translation with body fixation. Together, these oppositely directed shifts demonstrate that translations with world fixation were perceived longer than equivalent translations with body fixation, regardless of which translation was the reference.  Similarly, the shifts in \panelref{p3:fig2}{B} shows that world fixation translations were also perceived to be longer than free fixation movements and \panelref{p3:fig2}{C} shows that free fixation translations were perceived to be longer than body fixation translations. Note that \figref{p3:fig2} also shows effects on slope, which will be further discussed in the paragraph \nameref{p3:sec:precision}.

@@

    \caption{\panel{A} Actual (solid lines) and ideal (dashed lines) eye movement traces of one participant during world fixation (purple), body fixation (brown), and free fixation (black). All traces shown are for 10 \si{\centi\metre} reference movements. \panel{B} Normalised eye position for each participant (\textpm95\% confidence interval) at the end of translation interval (error bars) for world fixation (purple), body fixation (brown) and free fixation (blue). In addition, the average {\textpm}SE across all participants is shown. Zero indicates that the eyes remained stationary relative to the body, and one indicates that eye position was perfectly world-fixed.}
    \label{p3:fig3}
\end{figure}

@@

In order to relate psychophysical performance to eye movement behaviour we recorded and analysed eye movements during both intervals of every trial for all subjects. Exemplar eye traces for the 10 \si{\centi\metre} reference translation for the three fixation types are depicted in \panelref{p3:fig3}{A}. Fixation behaviour was quite accurate for both body fixations, where no eye movements were expected, and world fixation, where eye movement excursions of \siang{11} were expected, seemingly supported by  catch-up saccades. Under free fixation, the amount of eye movement was intermediate between body and world fixation and behaviour was more variable. A similar pattern was observed in all participants, as illustrated by the normalised eye movement data (see \nameref{p3:sec:methods}, and \panelref{p3:fig3}{B}).

@@

    \caption{Eye movement based prediction for the PSE plotted against the actual PSE.  A data point (symbol) is shown for each participant (symbol shape) and condition (symbol colour) pair, following the same colour scheme as in \figref{p3:fig2}. The identity line, corresponding to a perfect prediction, is shown in black.}
    \label{p3:fig4}
\end{figure}

@@

    \caption{Estimated eye movement contribution ($\alpha$) to the perception of self-motion (see \eqnref{p3:eq5}). Standard deviations are based on a bootstrap for each participant.}

@@

To quantify the role of eye movements in self-motion perception, we tested a linear model (see \nameref{p3:sec:model}) in which perceived translation is a weighted average of a vestibular estimate (equal to the actual translation) and an oculomotor estimate (equal to the normalised eye movement times the actual translation; \eqnref{p3:eq2}). This model contains a single free parameter ($\alpha$), which corresponds to the relative weight given to the oculomotor estimate. We fit this model to the two body versus world conditions and obtained the value of the oculomotor weight for every subject (\tabref{p3:tab2}). The average oculomotor weight is 0.25 \textpm0.12 (SD), indicating that the contribution of the eye movement signal to the self-motion estimate is about 25 percent. Note that participant 4, whose oculomotor weight is furthest from this mean ($\alpha = 0.06$), also shows a radically different eye movement gain during the free-fixation (see \panelref{p3:fig3}{B}). We then used these oculomotor weights along with the normalised eye movement values to predict the PSEs in the remaining four conditions according to \eqnref{p3:eq5}. The predicted PSEs are plotted against the actually observed PSEs in \figref{p3:fig4}. The positive correlation ($\rho = 0.78$, $p < 0.01$) between observed and predicted PSEs suggests that eye movements are indeed used in self-motion perception, even in the absence of a fixation point (i.e. during free fixation). Furthermore, the fact that data points generally cluster near the unity line shows that our simple model does reasonably well in predicting perceptual performance across subjects and conditions based on oculomotor weight and normalised eye movement magnitude only. This holds true even for subject 7 whose oculomotor weight (\tabref{p3:tab2}) was approximately double the average, yet whose data points remain close to the unity line.

@@

    \caption{Effect of difference movement amplitude between reference and probe interval (i.e. the bias) on response uncertainty ($\sigma$). A data point is shown for every participant and condition (symbol colour) pair, following the same colour scheme as in \figref{p3:fig2}. The dashed black line is the linear regression trend line.}

@@

 To quantitatively characterise the separate vestibular and eye movement contributions, we fit a single parameter model to the perceptual responses for the body versus world comparison conditions and validated this model independently by predicting the effects of eye movements on self-motion perception during free fixation conditions. This model takes into account subject specific oculomotor weight and eye movement patterns. Based on these inputs it accurately predicts the responses in the free fixation conditions. This demonstrates that extra-retinal eye movement signals are used as a cue in the perception of self-motion, contributing significantly to the self-motion percept with a weight of approximately 25 percent, even in the absence of optic flow.

It is surprising that an influence of eye movements can be observed even for body- stationary fixations, during which the stationary eye movement signal is clearly in conflict with the non-zero vestibular signal. While this demonstrates the strength of the assumption that fixation targets are world-stationary, it raises the question how reliable this assumption is. Simultaneous recording of angular head and eye movements during natural behaviour reveals that approximately 80 percent of eye movements can be classified as compensatory, i.e. eye movements directed opposite to head movement and therefore consistent with maintenance of world-fixed fixation \cite{einhauser2007}. Similarly, other studies have shown that world-stationary fixations are common for many every day activities, ranging from making a cup of tea \cite{hayhoe2014} to driving a car \cite{land1994}, to walking \cite{foulsham2011} and even reaching, where people tend to look at the source and destination of the object, but not at the hand \cite{flanagan2003}. Because world-stationary fixations are so common, the natural world statistics imply that self-motion and eye movements are highly correlated, thus making eye movements a fairly reliable cue for self-motion.

@@

Studies conducted to characterise vestibular-only sensitivity are often performed in complete darkness or with closed eyes \cite{grabherr2008,  macneilage2010b, macneilage2010a, roditi2012, valko2012, nesti2014}. However, the results of our free-fixation condition suggest that even under these circumstances, results could easily be influenced by vestibularly driven eye movements. Overall, we suggest that any study concerned with self-motion processing must consider the possible influence of eye movements.
